# pipeR Tutorial

```{r}
library(tidyverse)
library(pipeR)
```

## First Example 

```{r}
## bootstrap resampling over mpg

## traditional
sample_mtcars <- sample(mtcars$mpg, 10000, replace = TRUE)
density_mtcars <- density(sample_mtcars, kernel = "gaussian")
plot(density_mtcars, col = "red", main = "Density of MPG (bootstrap)")

## nested
plot(density(sample(mtcars$mpg, 10000, replace = TRUE), kernel = "gaussian"),
     col = "red", main = "Density of MPG (bootstrap)")

## pipe style 
mtcars$mpg %>>%
    sample(size = 10000, replace = TRUE) %>>%
    density(kernel = "gaussian") %>>%
    plot(col = "red", main = "Density of MPG (bootstrap)")

## pipe style with filtering
mtcars %>>%
    subset(mpg >= quantile(mpg, 0.05) &
           mpg <= quantile(mpg, 0.95)) %>>%
    ## pull mpg?
    (mpg) %>>%
    sample(size = 10000, replace = TRUE) %>>%
    density(kernel = "gaussian") %>>%
    plot(col = "red", main = "Density of MPG (bootstrap)")
```

## Pipe Operator 

The operator pipes the left hand side value to the right hand side expression which is evaluated according to its syntax. 

%>>% supports piping LHS values:
1. To the first argument of a function 
2. To . in an expression 
3. By formula to avoid ambiguity in symbol names
4. To an expression only for its side effect 
5. To an expression and save the intermediate results 
6. To a symbol in order to extract that element 

The behavior of the operator is fully determined by the syntax of the expression that follows the operator. 

## Pipe to First Argument 

```{r}
## original
set.seed(123)

summary(sample(diff(log(rnorm(100, mean = 10))), size = 10000, replace = TRUE))

## pipes 
rnorm(100, mean = 10) %>>%
    log %>>%
    diff %>>%
    sample(size = 10000, replace = TRUE) %>>%
    summary

## multiple outputs
mtcars %>>%
    (mpg) %>>%
    plot(col = "red",
         main = sprintf("Number of Points: %d", length(.)))

## nesting in function calls
mtcars %>>%
    subset(mpg <= mpg %>>% quantile(0.95), c(mpg, wt)) %>>%
    summary
```

## Pipe to Dot 

```{r}
## second arg example
mtcars %>>%
    {lm(mpg ~ cyl + wt, data = .)}

mtcars %>>%
    (lm(mpg ~ cyl + wt, data = .))
```

The difference between {} and () above is

1. {} accepts more than one expression within the braces and its value is determined by the last one. () accepts only 1 expression.
2. {} has only 1 feature: pipe to . in the enclosed expression. () has more features (coming soon)

```{r}
## demonstrate both
mtcars %>>% {
    model <- lm(mpg ~ wt + cyl, data = .)
    summ <- summary(model)
    summ[c("r.squared", "adj.r.squared")]
}

## returns last expression unless interrupted by a return
mtcars %>>%
    lm(formula = mpg ~ wt + cyl) %>>%
    summary %>>% {
        if (.$r.squared >= 0.8) return("Model A")
        else cat("Model B should be considered.\n"); return("Model B")
    }

## side effects
mtcars %>>% {
    par(mfrow = c(1, 2))
    .$mpg %>>% plot(col = "red", main = "mtcars (mpg)")
    .$mpg %>>% hist(main = "distribution")
}
    
## nested dots are determined by the working operator for that dot symbol
mtcars %>>% {
    par(mfrow = c(1, 2))
    .$mpg %>>% plot(col = "red", main = sprintf("mtcars (mpg: %d)", length(.)))
    .$mpg %>>% hist(main = "distribution")
}
```


### Creating Closure 

If a lambda function is enclosed, the pipeline will result in a closure, a function returns by a function (%>>%) which can be used more times. 

```{r}
## create a closure that takes 1 element
mtcars %>>%
    (mpg) %>>%
    sample(size = 10000, replace = TRUE) %>>%
    (function(kernel) {
        . %>>%
            density(kernel = kernel) %>>%
            plot(main = sprintf("%s Kernel", kernel))
    }) -> density_plot

## return plots
par(mfrow = c(1, 3))
map(c("gaussian", "rectangular", "triangular"), ~ density_plot(.x))

## %>>% will pipe the left value to the first arg of the closure
1:10 %>>% (function(x, pow) x^pow)(2)

mtcars %>>%
    lm(formula = mpg ~ wt + cyl) %>>%
    (function(model, warn_level) {
        if (summary(model)$r.squared < warn_level)
            warning("R Squared is too low", call. = FALSE)
        model
    })(0.9) %>>%
     coef
```

# Pipe by Formula 

```{r}
## something like this could be confusing due to the double dot
mtcars %>>%
    (lm(mpg ~ ., data = .))


## we can use a lambda function to represent the value being piped
mtcars %>>%
    (df ~ lm(mpg ~ ., data = df))

## example mixing first arg piping and piping by formula
mtcars %>>%
    subset(select = c(mpg, wt, cyl)) %>>%
    (x ~ lm(mpg ~ ., data = x))

## example with brackets
mtcars %>>%
    subset(select = c(mpg, wt, cyl)) %>>%
    (x ~ {
        summ <- lm(mpg ~ ., x) %>>% summary()
        list(n = nrow(x),
             r.squared = summ$r.squared)})
```

## Pipe for Side Effect 

%>>% provides a syntax to support side effect piping. The syntax is called a one-sided formula that starts with ~ to indicate that the right hand side expression will be evaluated only for its side effect. Its values will be ignored, and the input value will be returned instead. 

```{r}
mtcars %>>%
    subset(mpg >= quantile(mpg, 0.05) &
           mpg <= quantile(mpg, 0.95),
           c(mpg, cyl, wt)) %>>%
    ## cat returns NULL
    (~ cat("Rows:", nrow(.), "\n")) %>>%
    summary()
```

The (~ f(.)) syntax tells the operator that f(.) on the right of ~ should only be evaluated for its side effect with . representing the input value, and the input value, i.e. the filtered mtcars, should continue in the pipeline and be passed to summary. The design of the ~ expression syntax for side effects makes the one-sided formula a branch of the mainstream pipeline. 

```{r}
## example of plotting in the middle a pipeline as a side effect
mtcars %>>%
    subset(mpg >= quantile(mpg, 0.05) &
           mpg <= quantile(mpg, 0.95),
           c(mpg, cyl, wt)) %>>%
    (~ plot(mpg ~ wt, data = .)) %>>%
    lm(formula = mpg ~ wt) %>>%
    summary()
```

### Lambda Expression 

If you feel ambiguous using ., you can write a formula after ~ 

```{r}
mtcars %>>%
    subset(mpg >= quantile(mpg, 0.05) &
           mpg <= quantile(mpg, 0.95),
           c(mpg, cyl, wt)) %>>%
    (~ filtered ~ cat("Rows:", nrow(filtered), "\n")) %>>%
    summary()
```

### String 

```{r}
## print a string in the middle of a pipeline
mtcars %>>%
    "Running linear model" %>>%
    lm(formula = mpg ~ wt + cyl) %>>%
    "Summarizing the model" %>>%
    summary %>>%
    coef

## print strings to indicate working progress of the pipeline
diamonds %>>%
    "Filtering Data" %>>%
    filter(carat >= quantile(carat, 0.05) &
           carat <= quantile(carat, 0.95)) %>>%
    (~ Sys.sleep(2)) %>>%
    "Selecting Variables" %>>%
    select(carat, price, depth) %>>%
    (~ Sys.sleep(2)) %>>%
    "Estimating Linear Model" %>>%
    lm(formula = carat ~ price + depth) %>>%
    (~ Sys.sleep(2)) %>>%
    "Summarizing Model" %>>%
    (~ Sys.sleep(2)) %>>%
    summary()
```

### Question Mark 

An easier way to print the intermediate value is to use the (? expr) syntax 

```{r}
## print intermediate values
mtcars %>>%
    subset(vs == 1, c(mpg, cyl, wt)) %>>%
    (? nrow(.)) %>>%
    (? data ~ ncol(data)) %>>%
    summary()

## reduce ambiguity with prefix strings
mtcars %>>%
    ("Raw Data Rows:" ? nrow(.)) %>>%
    subset(vs == 1, c(mpg, cyl, wt)) %>>%
    ("Filtered Data Rows:" ? nrow(.)) %>>%
    lm(formula = mpg ~ cyl + wt)
```

### Stopping 

Side effects are usually used on purpose not to influence the mainstream pipeline or workflow. In some cases however, such influence can be useful. An example is stopping. 

```{r}
## build an lm. Before making preds, check if R^2 > .5
mtcars %>>%
    subset(vs == 1, c(mpg, cyl, wt)) %>>%
    lm(formula = mpg ~ cyl + wt) %>>%
    (~ stopifnot(summary(.)$r.squared >= 0.5)) %>>%
    predict(newdata = list(cyl = 4.45, wt = 3.0))

## set threshold to 0.8
mtcars %>>%
    subset(vs == 1, c(mpg, cyl, wt)) %>>%
    lm(formula = mpg ~ cyl + wt) %>>%
    (~ stopifnot(summary(.)$r.squared >= 0.8)) %>>%
    predict(newdata = list(cyl = 4.45, wt = 3.0))
```

### Debugging 

Another way to use a side effect is debugging. 

```{r}
## go into debug mode in the middle of a pipeline
mtcars %>>%
    subset(vs == 1, c(mpg, cyl, wt)) %>>%
    lm(formula = mpg ~ cyl + wt) %>>%
    summary() %>>%
    (~ browser(expr = .$r.squared < 0.8))
```

## Pipe with Assignment 

Perhaps the most important side effect of all is assignment. 

```{r}
## assign intermediate steps
mtcars %>>%
    subset(select = c(mpg, wt, cyl)) %>>%
    ## assign subsetted mtcars
    (~ sub_mtcars) %>>%
    lm(formula = mpg ~ wt + cyl) %>>%
    (~ lm_cars) %>>%
    summary()

sub_mtcars
lm_cars

## assign the intermediate value under some transformation
mtcars %>>%
    subset(select = c(mpg, wt, cyl)) %>>%
    ## side effect assignment
    (~ summ <- summary(.)) %>>%
    lm(formula = mpg ~ wt + cyl)

## assignment using named lambda vars
mtcars %>>%
    subset(select = c(mpg, wt, cyl)) %>>%
    (~ summ <- df ~ summary(df)) %>>%
    lm(formula = mpg ~ wt + cyl)
```

What if one wants the result not only to be assigned to a symbol, but also to continue the flow to the next expression? 

Two methods meet the demand: 
1. Use (~ symbol) after the expression for assignment 
2. Use (symbol = expression) to assign the value of expression to symbol

```{r}
## pipe and assignment. Pipes values to lm, and assigns lm to model
mtcars %>>%
    subset(select = c(mpg, wt, cyl)) %>>%
    ## side effect assignment
    (~ summ <- df ~ summary(df)) %>>%
    ## pipe and assign
    (model <- lm(mpg ~ wt + cyl, data = .))

## use -> for assignment
mtcars %>>%
    (~ summary(.) -> summ)

mtcars %>>%
    ## side effect assignment
    (~ summary(.) -> summ) %>>%
    (lm(formula = mpg ~ wt + cyl, data = .) -> lm_mtcars) %>>%
    summary()

## piped assigment operators support subset and element assignment
results <- list()

mtcars %>>%
    lm(formula = mpg ~ wt + cyl) %>>%
    (~ results$mtcars <- . %>>% summary %>>% (r.squared))

iris %>>%
    lm(formula = Sepal.Length ~ Sepal.Width) %>>%
    (~ results$iris <- . %>>% summary %>>% (r.squared))
    
results

## set names and others
numbers <- 1:5

letters %>>%
 sample(length(numbers)) %>>%
 (~ . -> names(numbers))

numbers
```

## Extract Element 

In some cases, one might need to extract an element from an object. The syntax x %>>% (y) indicates to extract the element named y from object x where y must be a valid symbol name, and x can be a vector, list, environment, or anything else that supports [[]]. 

```{r}
## example
mtcars %>>%
    (lm(mpg ~ wt + cyl, data = .)) %>>%
    summary %>>%
    (r.squared)

## examples for vector, list, and environment
list(a = 1, b = 2) %>>% (a)
c(a = 1, b = 2) %>>% (a)

env <- new.env()
env$a <- 1; env$b <- 2
env %>>% (a)

## S4 object example

## create a data generating process
set.seed(0)
x <- cumsum(rnorm(150))
y <- 2 * x + rnorm(150) * 0.5
matplot(cbind(x, y), type = "l")

## perform an ADF test to test the cointegration relationship between two vectors on the residues of a linear model
lm(y ~ x + 0) %>>%
    residuals %>>%
    ## s4 object
    fUnitRoots::adfTest(type = "nc") %>>%
    ## save test object to env
    (~ testobj) %>>%
    ## what class is it?
    (? class(.)) %>>%
    ## get testobj@test
    (test) %>>%
    ## get testobj@test$statistic
    (statistic)
```

Since the syntax only accepts symbol name, to evaluate an expression with the piped object being the frame if it is a list or environment, using with() or within() can be helpful

```{r}
list(a = 1, b = 2) %>>%
    with(a + 2 * b)
```

# Pipe Function 

todo 

## Pipe to First Argument 

## Dot Function

## Subsetting 

# Pipeline 

todo

## Argument-based Pipeline 

## Expression-based Pipeline 

# Examples

## dplyr

```{r}
# load data
library(hflights)
data(hflights)
hflights %<>% as_tibble()
```

If we divide Distance by ActualElapsedTime we can get the actual flight speed. 

```{r}
hflights %>>%
    filter(Cancelled == 0) %>>%
    mutate(speed = Distance / ActualElapsedTime) %>>%
    ## save to hflights2
    (~ hflights2) %>>%
    group_by(UniqueCarrier) %>>%
    summarize(mean_speed = mean(speed, na.rm = TRUE)) %>>%
    arrange(desc(mean_speed)) %>>%
    with(barplot(mean_speed, names.arg = UniqueCarrier, main = "Average Flight Speed"))

## account for destinations
hflights2 %>>%
    dcast(Year + Month + DayofMonth ~ UniqueCarrier,
          value.var = "speed", fun.aggregate = mean) %>>%
    subset(select = c(-1, -2, -3)) %>>%
    ## save wide format data 
    (~ hflights3) %>>%
    list.sort(desc(median(., na.rm = TRUE))) %>>%
    boxplot(main = "Boxplot of Flight Speed Sorted by Median")

```

## rlist 

## rvest 

