# rlist Tutorial 

```{r}
library(rlist)
library(pipeR)
```

## Quick Overview

```{r}
url <- "http://renkun-ken.github.io/rlist-tutorial/data/people.json"

people <- list.load(url)

## the tutorial really needs the json file unfortunately
people %>>%
    list.filter(Expertise$R >= 1 & Expertise$Python >= 1) %>>%
    list.class(Interests) %>>%
    list.sort(-length(.)) %>>%
    list.take(3) %>>%
    list.map(. %>>% list.table(Age))

## example in which we also use map from purrr. Works on tibbles
mtcars %>%
    map(., ~ summary(.x) %>% broom::tidy()) %>%
    list.filter(median > 190)
```

# Features 

## Mapping

### list.map

```{r}
## list. load is designed for loading data from a given data source
people <- list.load(url)
str(people)

## pull all the names
list.map(people, Name)

## another one
list.map(people, Age) %>>%
    list.filter(between(., 25, 27))

## summarization
list.map(people, sum(as.numeric(Expertise)))

## return multiple values
list.map(people, list(Age = Age,
                      Range = range(Expertise)))

## in some cases, we need to refer to the item itself, its index in the list, or even its name
## . is the item itself, .i represents its index, .name represents its name
nums <- c(a = 1, b = 2, c = 3)

## itself
list.map(nums, . + 1)

## index
list.map(nums, .i)

## name
list.map(nums, .name)
```

### list.mapv

If we wish to return a vector rather than a list, we can use list.mapv, which basically calls unlist to the list resulted from list.map()

```{r}
list.mapv(people, Age)

list.mapv(people, sum(as.numeric(Expertise))) %>%
    list.filter(. > 10)
```

### list.select 

This is as expected. This can be done with list.map, but this is nicer syntax

```{r}
list.select(people, Name, Age)

list.map(people, list(Name, Age))

list.select(people, Name, Age, nlang = length(Expertise))
```

### list.iter 

sometimes we don't really need the result of a mapping but its side effects. This is essentially walk 

```{r}
list.iter(people, cat(Name, ":", Age, "\n"))
```

### list.maps 

All the previous functions work with a single list. list.maps works with multiple lists -- each of which is represented by a user-defined symbol at the function call 

```{r}
l1 <- list(p1 = list(x = 1, y = 2),
           p2 = list(x = 3, y = 4),
           p3 = list(x = 1, y = 2))

l2 <- list(2, 3, 5)

list.maps(a$x * b + a$y,
          a = l1, b = l2)

## the function doesn't require names be supplied
list.maps(..1$x * ..2 + ..1$y, l1, l2)
```

## Filtering 

### list.filter 

As opposed to list mapping which evaluates an expression given each list element, list filtering evaluates an expression to decide whether to include the entire element in the results. 

```{r}
list.filter(people, Age >= 25)

## get names of people who are less than 25
people %>>%
    list.filter(Age >= 25) %>>%
    list.mapv(Name)

## get the names of those interested in music
people %>%
    list.filter("music" %in% Interests) %>%
    list.mapv(Name)

## get peole who have been using programming languages for at least 3 years on average
people %>%
    list.filter(mean(as.numeric(Expertise)) >= 3) %>%
    list.mapv(Name)

## meta symbols like ., .i, .name can also be used
people %>%
    list.filter(.i %% 2 == 0) %>%
    list.mapv(Name)
```

### list.find 

list.find avoids searching the entire list and returns a certain number of items found 

```{r}
people %>%
    list.find(Age >= 25, 1) %>%
    list.mapv(Name)
```

### list.findi 

similar to list.find, but it returns the index 

```{r}
list.findi(people, Age >= 23, 2)
```

### list.first, list.last 

These are used to find the first and last element that meet certain conditions 

```{r}
list.first(people, Age >= 23)

list.last(people, Age >= 23)

people %>%
    list.sort(Name) %>%
    list.last(Age >= 23)

## they also return the first / last element of a vector if the predicate is missing
list.first(1:10)
list.last(1:10)
```

### list.take 

list.take takes at most a given number of elements from a list. 

```{r}
list.take(1:10, 3)

list.take(1:5, 8)
```

### list.skip 

skips a given number of elements in the list and returns the rest as results

```{r}
list.skip(1:10, 3)

list.skip(1:5, 8)
```

### list.takeWhile 

keeps taking elements until a condition holds true 

```{r}
people %>%
    ## list.map(list(., Expertise$R)) %>% 
    list.takeWhile(Expertise$R >= 2) %>%
    list.map(list(name = Name,
                  R = Expertise$R))
```
